---
title: "The Effect of Economic Events on Votes for the President"
subtitle: "Resource Economics 312"
author: "Scott Cohn and Samuel Hostetter"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      incremental: true

      
---
```{r setup, include=FALSE, cache=TRUE}
knitr::opts_chunk$set(echo = FALSE, fig.align = "center")

# Libraries
library(rlang)
library(readxl)
library(dplyr)
library(randomForest)
library(car)
library(ggplot2)
library(ggthemes)
library(sjPlot)
library(forecast)
library(xtable)
library(broom)
library(knitr)
library(lmtest)
library(het.test)
library(plotly)
library(ggfortify)
library(party)
library(dynlm)
library(stargazer)
library(kableExtra)
library(knitrBootstrap)
library(readr)
library(magick)
library(dygraphs)
library(magrittr) 


# Import

ModelData <- read_excel("data.xlsx")
D_Outcome <- read_excel("data.xlsx", sheet = "Actual Value")
FairVersusActual <- readxl::read_excel("data.xlsx", sheet = "Fair_Versus_Actual_Pres")

# Trim and Clean
WorkSet <- ModelData[11:35, 1:10]
#WorkSet <- WorkSet[, -WorkSet$rfpred]
names(WorkSet)[names(WorkSet) == 't'] <- 'Year'


dplyr::inner_join(WorkSet, D_Outcome, by = "Year")

WorkSet <-
  data.frame(lapply(WorkSet, function(x)
    as.numeric(as.character(x))))

# Convert VP and actualVP to percentages
WorkSet$VP <- (WorkSet$VP / 100)
WorkSet$actualVP <- (WorkSet$actualVP / 100)

# Define WorkSet
explore_data <- WorkSet[ ,c("Year", "actualVP", "I", "DPER", "DUR", "WAR", "G", "P", "Z")]

# Models
full_model <- lm(actualVP ~ I + DPER + DUR + WAR + G + P + Z, data = explore_data)
interaction_model <- lm(actualVP ~ I + DPER + DUR + WAR + G*I + P*I + Z*I , data = explore_data)
model_Fair <- lm(actualVP ~ I + DPER + DUR + WAR + G:I + P:I + Z:I , data = explore_data)
Fair_NoWar <- lm(actualVP ~ I + DPER + DUR + G:I + P:I + Z:I , data = explore_data)
```
class: incremental
# Literature Review

- Anthony Downs
  + Establishes a disconnect between voter desires and government desire
  
- Gerald H. Kramer
  + Focused on votes for the House of Representatives and Congress
  + Argued effect of presidential elections is much more influenced by non-economic events
  
- George J. Stigler
  + Believes voters think about many confounding factors when they vote
  + Voters are "self-interested and well-informed"

---
class: middle, center

# The Model 

$$V_t = \alpha_1 + \alpha_2 G_t * I_t + \alpha_3 P_t * I_t + \alpha_4 Z_t * I_t + \alpha_5 {DPER_t} + \alpha_6 {DUR_t} +\alpha_7 I_t + \alpha_8 {WAR_t} + \mu_t$$
---

```{r VarTable, echo=FALSE, message=FALSE, warning=FALSE, cache=TRUE}
VarDef <- read_excel("data.xlsx", sheet = "VarDef")
kable(VarDef,
      format = "html",
      #table.attr = "style = 'width:90%;'",
      booktabs = T
      ) %>%
  #column_spec(2, width = "40em") %>%
  kableExtra::kable_styling(bootstrap_options = c("striped", "condensed", position = "center"), font_size = 17)# %>%
 # kable_as_image("vardef_table")

```
---
# Ray Fair

- Probably wrote own of your textbooks (Case, Fair, Oster)

- Derived first model in Fair (1978)

- Not much changed until 1992

---
class:middle

- How do economic events affect the voting behavior?

- Two purposes
  + Present general model to incorporate theory of the time
  + Model should allow testing of these theories against each other

---

```{r FairModel, echo=FALSE, message=FALSE, warning=FALSE}
library(magrittr)
dygraphs::dygraph(FairVersusActual,
        main = "Model Prediction vs Actual Outcome",
        ylab = "Vote Percentage",
        xlab = "Year",
        width = 750) %>% 
  dygraphs::dySeries("Fair_Prediction", label = "Prediction") %>%
  dygraphs::dySeries("Actual_Vote_Share", label = "Actual Vote") %>%
  # dyGroup(c("Fair_Prediction", "Actual_Vote_Share"), drawPoints = TRUE) %>% 
  dygraphs::dyOptions(stackedGraph = FALSE) %>%
  dygraphs::dyRangeSelector(height = 10) %>% 
  dygraphs::dyAxis("x", drawGrid = FALSE) %>% 
  dygraphs::dyLegend(width = 500) %>% 
  dygraphs::dyEvent("1964", "LBJ", labelLoc = "bottom") %>%
  dygraphs::dyEvent("1992", "Ross Perot", labelLoc = "bottom")
```

---
class:middle, center

$V_t = \alpha_1 + \alpha_2 G_t + \alpha_3 P_t + \alpha_4 Z_t + \alpha_5 {DPER_t} + \alpha_6 {DUR_t} +\alpha_7 I_t + \alpha_8 {WAR_t} + \mu_t$

---

```{r echo=FALSE, cache=FALSE, fig.align="center"}
library(sjPlot)
#summary(full_model)
sjt.lm(full_model,
       show.ci = FALSE,
       show.fstat = TRUE,
       show.se = TRUE,
       show.est = TRUE
       )
```

---
class: middle, center

$V_t = \alpha_1 + \alpha_2 G_t * I_t + \alpha_3 P_t * I_t + \alpha_4 Z_t * I_t + \alpha_5 {DPER_t} + \alpha_6 {DUR_t} +\alpha_7 I_t + \alpha_8 {WAR_t} + \mu_t$

---

```{r echo=FALSE, cache=FALSE, fig.align="center"}
#summary(interaction_model)
sjt.lm(interaction_model,
       show.ci = FALSE,
       show.fstat = TRUE,
       show.se = TRUE,
       show.est = TRUE
       )
```

---
class: middle, center

$V_t = \alpha_1 + \alpha_2 G_t * I_t + \alpha_3 P_t * I_t + \alpha_4 Z_t * I_t + \alpha_5 {DPER_t} + \alpha_6 {DUR_t} +\alpha_7 I_t + \alpha_8 {WAR_t} + \mu_t$

---

```{r echo=FALSE, cache=FALSE, fig.align="center"}
#summary(model_Fair)
sjt.lm(model_Fair,
       show.ci = FALSE,
       show.fstat = TRUE,
       show.se = TRUE,
       show.est = TRUE
       )
```

---
# Multicollinearity

$$VIF_k = \frac{1}{1 - R^2_k}$$

```{r VIF, message=FALSE, warning=FALSE, echo=FALSE, cache=TRUE}
tab <- broom::tidy(vif(model_Fair))
kable(tab, 
    format = "latex",
    booktabs = T,
    align = "c",
    digits = 2,
    col.names=c("regressor","VIF")) %>% 
  kableExtra::kable_styling(font_size = 8)  %>% 
  kable_as_image("VIF_Table")
```
---
# Heteroskedasticity

$$\chi^2 = n\times R^2 \sim \chi^2_{(N-1)}$$
```{r echo=FALSE, cache=TRUE}
kable(tidy(bptest(model_Fair)),
    format = "latex",
    booktabs = T,
    align = "c",
    digits = 3
    )  %>% 
  kableExtra::kable_styling(font_size = 8) %>% 
  kable_as_image("Heterosk_Table")
```
---
# Autocorrelation

$$d = \frac{\sum_{t=2}^{T}(e_t-e_{t-1})^2}{\sum_{t=1}^{T}e^2_t}$$
```{r DurbinWatson, echo=FALSE, message=FALSE, warning=FALSE, cache=TRUE}
DW <- dwtest(model_Fair)
kable(tidy(DW),
    format = "latex",
    booktabs = T,
    align = "c",
    digits = 3
    )  %>% 
  kableExtra::kable_styling(font_size = 8) %>% 
  kable_as_image("Autocorr_Table")

```
---
# Forecasts

```{r echo=FALSE, cache=TRUE}
Fair2014Forecast <- read_excel("data.xlsx", sheet = "Fair2014Forecast")
kable(Fair2014Forecast,
    format = "latex",
    booktabs = T,
    digits = 2
    )  %>% 
  footnote(general = "2016 Forecast") %>% 
  kable_styling("striped", font_size = 8)  %>% 
  kable_as_image("2016_Forecast")

Forecast2020 <- read_excel("data.xlsx", sheet = "Forecast")
kable(Forecast2020,
    format = "latex",
    booktabs = T,
    digits = 2
    )  %>% 
  footnote(general = "2020 Forecast") %>% 
  kable_styling("striped", font_size = 8)  %>% 
  kable_as_image("2020_Forecase")
```
---
class: center, middle

# What's Next?


